A Generalized Sphere Theorem for\\Positively Curved Combinatorial 3-Manifolds

### 2013-06-25 paper/paper.tex
more journal specific stuff added

@@

\nocite{*}

\title{A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds}

\author{Aaron Trout \\ Department of Mathematics \\
Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA \and
Vadas Gintautas\\ Department of Physics \\
Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA}

@@

\begin{abstract}
The generalized sphere theorem of Grove and Shiohama is an important result in the differential geometry of positively curved manifolds. It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere. In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds. That is, those with at most five tetrahedra incident at each edge, or equivalently, those with an angle deficit at each edge in the standard piecewise-linear metric. Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem. Here, we use a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan to prove a corresponding {\em discrete} generalized sphere theorem.
\end{abstract}

>>>>

\title{A Generalized Sphere Theorem for\\Positively Curved Combinatorial 3-Manifolds}

\author{Aaron D. Trout\\
\small Department of Mathematics\\[-0.8ex]
\small Chatham University\\[-0.8ex] 
\small Pittsburgh PA 15232,  U.S.A.\\
\small\tt atrout@chatham.edu\\
\and
Vadas Gintautas\\
\small Department of Physics\\[-0.8ex]
\small Chatham University\\[-0.8ex] 
\small Pittsburgh PA 15232,  U.S.A.\\
\small\tt vgintautas@chatham.edu}

\date{\dateline{June 25, 2013}{XX}\\
\small Mathematics Subject Classifications: 52C99, 53A99, 57M99}

@@

@@

\begin{abstract}
The generalized sphere theorem of Grove and Shiohama is an important result in the differential geometry of positively curved manifolds. It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere. In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds. That is, those with at most five tetrahedra incident at each edge, or equivalently, those with an angle deficit at each edge in the standard piecewise-linear metric. Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem. Here, we use a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan to prove a corresponding {\em discrete} generalized sphere theorem.

\bigskip\noindent \textbf{Keywords:} discrete differential geometry; generalized sphere theorem; piecewise-linear geometry; positively curved combinatorial manifold
\end{abstract}
